% Thesis Declaration ---------------------------------------------------
\begin{declaration} 
I declare that the work presented in this thesis is my own, other than where clearly stated. The work done by others has been explicitly referenced accordingly, this includes passages of text, figures, plots and images. Some of this work has been the result of collaborative work and all participants have been identified within this thesis. 

%I can confirm that the work presented in this thesis has not been submitted for another qualification to this or any other university.

Chapters 1 and 2 describe the history, background and theoretical motivation for the subject of this thesis. These chapters are then a summary of the vast amount of work conducted by others, gathered from scientific journals, papers, documents and published books, which have been referenced accorindinly.

Chapters 3 and 4 then focus on the LAGUNA-LBNO project and its aims, development and work, with the latter chapter focusing on the near detector within LAGUNA-LBNO. The work has been conducted within the collaboration, with the published work and internal documents referenced when applicable. Chapter 4 involves a smaller collaborative effort within LAGUNA-LBNO, in which I have contributed significantly.

Chapter 5 represents the analysis work regarding the near detector within LAGUNA-LBNO. This chapter describes the simulation framework, the detector geometry design and modelling, simulation studies and data analysis. I made significant contributions to the software framework as one of two main developers and I also made major contributions to the detector design. The rest of the work in the chapter, involving the modelling and analysis, represents my effort and work. An introductory background is presented at the start of this chapter motivating Monte Carlo simulations and describes the third party software used, all of which is referenced accordingly. This work has been presented to the collaboration internally, represented in internal documents and presented outside the collaboration to members of the scientific community.

Chapter 6 presents the MODES-SNM project in a similar fashion to chapter 3, where it represents the work of the MODES-SNM collaboration. The work shown in this chapter is a summary of published work and internal documents from members of the collaboration. These have been referenced accordingly and members are noted in this chapter.

Chapter 7 describes the detectors and the integration of the MODES-SNM system. Much of this work was conducted at the University of Liverpool, where I made significant contributions to the design, construction and integration of the system. The majority of the work conducted by myself is represented by the simulation studies and the integration.

Chapter 8 describes the MODES-SNM system demonstrations in various locations across Europe in a real life environment. For all the demonstrations described in the chapter I was only present for one, the demonstration at London Heathrow Airport, which I was significantly involved in the majority of the data collection and the demonstration report. Other demonstrations described in this chapter are based on shared collaboration documents and files. 

Chapter 9 forms the MODES-SNM analysis chapter, which presents my analysis framework and implementation. The data used was collected by other members of the MODES-SNM collaboration and was shared within the group. This has been referenced. The initial part of the chapter presents the work conducted by other members of the MODES-SNM collaboration used to motivate the remaining part of the chapter and is necessary for the analysis. This first section is based on internal documents within the collaboration and is explicitly stated so in the chapter, referenced accordingly.

\end{declaration}

% ----------------------------------------------------------------------

